---
title: "Initial Korean covid 19"
author: "Kimberly Aldana"
date: "June 21, 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?



# Methods

  We will analyze the sex of the patients and the results if they are alive or dead. I will use the variables in my collection of Cat~cat. This is a type of problem that we can use the helper file and we will copy into our project to get the answer if dying is independent of sex. 


## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

 The data in this bar plot allows us to see the results of the alive and dead people. We can see that twice as many males died more than females. The disease are killing more males than females according to the data investigated from South Korea 
  

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, kRdata)
rowPerc(table1)
colPerc(table1)
```

 I can see that the people that are still alive, 56 of females and 36 of males. The number of dead people were 44 females and 64 males. The percentage was 98.7% of females remained alive while only 1.4% of total females died. Only 97% of males remained of alive while 3% of males died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

 In the group of 1831 were alive and 1411, in all more female are alive than male. The p value is 1 out of a thousand in the Chi-Square. It is a small chance to get a male and female dying. In the fisher test, 9 out of 10,000, the are about the same p value. You can also see the odds of female against the odds of male dying. We want to know that the odds this is happening. Women have a 2.3% higher chance of surviving COVID-19 than males in the South Korea investigation.



## Conclusions

 In today's lesson I learned to do a ggplot with the helper's file. I changed the class for result and name the project kRdata. I was also able to create a numerical graph in % and counting. I also ran a chi-square test and a fisher's test and both are equal in p value. The final result is: that we want to know what the odds are of this happening. All of the males are the ones who has a higher chance of death vs woman during COVID-19 in this South Korea study. We can conclude that men are not likely to die equally as of woman; since woman have more of the advantage of survival than males.





